---
title: "UMAP demo"
output: html_notebook
---

# Loads UMAP
```{r}
library(umap)
library(ggplot2)
library(ggthemes)
```

# Loads Broad cancer dependency map dataset and Travers et al. 2015
# common core gene essentiality dataset. Note that I lightly processed
# the Depmap data - the most current version can always be found at 
# https://depmap.org/portal/
```{r}
depmap <- read.csv("~/../Desktop/depmap_formatted_1_24_19.tsv", sep = "\t")
essentials <- read.csv("~/../Desktop/traver_core_essential_2015.txt", sep = "\t")
```

# Appends essential gene info to depmap data and stores
# indices of columns with numerical data
```{r}
essentials <- essentials[,1]
depmap <- data.frame(depmap)
depmap$essential <- ifelse(rownames(depmap) %in% essentials, 
                           "Essential", "Non-essential")
data_cols <- !(colnames(depmap) %in% c("essential"))
```

# Runs UMAP on Depmap dataset and plots with colors according to
# gene essentiality
```{r}
umap_result <- umap(depmap[,data_cols])
umap_data <- data.frame(umap_result$layout)
colnames(umap_data) <- c("UMAP1", "UMAP2")
umap_data$essential <- depmap$essential
ggplot(umap_data, aes(x = UMAP1, y = UMAP2, color = essential)) +
  geom_point(alpha = 0.7) +
  scale_color_manual(values = c("Red", "Gray")) +
  labs(color = "Gene Essentiality") +
  theme_tufte(base_size = 14)
```